---
title: "R Notebook"
output: html_notebook
---

#Libraries
```{r}
# Version info: R 4.2.2, Biobase 2.58.0, GEOquery 2.66.0, limma 3.54.0, DESeq2 1.38.3
################################################################
#   Differential expression analysis with limma
library(GEOquery)
library(limma)
library(umap)
```

# load series and platform data from GEO
```{r}
gset <- getGEO("GSE60740", GSEMatrix =TRUE, AnnotGPL=FALSE)
if (length(gset) > 1) idx <- grep("GPL17811", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]
```
# make proper column names to match toptable 
```{r}
fvarLabels(gset) <- make.names(fvarLabels(gset))
```

# group membership for all samples
```{r}
gsms <- "0000000000000011111112233444455"
sml <- strsplit(gsms, split="")[[1]]
```

# assign samples to groups and set up design matrix
```{r}
gs <- factor(sml)
groups <- make.names(c("tumor_CIC-DUX4","tumor_EWSR-NFATc2","hMSC_EWSR-NFATc2", "hMSC_Ctrl","IB120_shCIC-DUX4", "IB120_shCtrl" ))
levels(gs) <- groups
gset$group <- gs
design <- model.matrix(~group + 0, gset)
colnames(design) <- levels(gs)

gset <- gset[complete.cases(exprs(gset)), ] # skip missing values

fit <- lmFit(gset, design)  # fit linear model
```

# set up contrasts of interest and recalculate model coefficients
```{r}
cts <- paste(groups[2], groups[4], sep="-")
cont.matrix <- makeContrasts(contrasts=cts, levels=design)
fit2 <- contrasts.fit(fit, cont.matrix)

# compute statistics and table of top significant genes
fit2 <- eBayes(fit2, 0.01)
tT <- topTable(fit2, adjust="fdr", sort.by="B", number=250)
```

```{r}
require(org.Hs.eg.db)
tT$SYMBOL <- AnnotationDbi::select(org.Hs.eg.db, 
       keys = as.character(tT$ENTREZ_GENE_ID),
       columns = c("ENTREZID", "SYMBOL"),
       keytype = "ENTREZID")$SYMBOL
```
```{r}
EWS.NFATc2 <- tT %>% top_n(100, logFC) %>% pull(SYMBOL)
```


# log2 transformation
```{r}
ex <- exprs(gset)
qx <- as.numeric(quantile(ex, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
LogC <- (qx[5] > 100) ||
          (qx[6]-qx[1] > 50 && qx[2] > 0)
if (LogC) { ex[which(ex <= 0)] <- NaN
  exprs(gset) <- log2(ex) }
```

# General expression data analysis
```{r}
ex <- exprs(gset)
```


# box-and-whisker plot
```{r}
ord <- order(gs)  # order samples by group
palette(c("#1B9E77", "#7570B3", "#E7298A", "#E6AB02", "#D95F02",
          "#66A61E", "#A6761D", "#B32424", "#B324B3", "#666666"))
par(mar=c(7,4,2,1))
title <- paste ("GSE60740", "/", annotation(gset), sep ="")
boxplot(ex[,ord], boxwex=0.6, notch=T, main=title, outline=FALSE, las=2, col=gs[ord])
legend("topleft", groups, fill=palette(), bty="n")
```


# expression value distribution
```{r}
par(mar=c(4,4,2,1))
title <- paste ("GSE60740", "/", annotation(gset), " value distribution", sep ="")
plotDensities(ex, group=gs, main=title, legend ="topright")
```


# UMAP plot (dimensionality reduction)
```{r}
ex <- na.omit(ex) # eliminate rows with NAs
ex <- ex[!duplicated(ex), ]  # remove duplicates
ump <- umap(t(ex), n_neighbors = 9, random_state = 123)
par(mar=c(3,3,2,6), xpd=TRUE)
plot(ump$layout, main="UMAP plot, nbrs=9", xlab="", ylab="", col=gs, pch=20, cex=1.5)
legend("topright", inset=c(-0.15,0), legend=levels(gs), pch=20,
col=1:nlevels(gs), title="Group", pt.cex=1.5)
library("maptools")  # point labels without overlaps
pointLabel(ump$layout, labels = rownames(ump$layout), method="SANN", cex=0.6)
```


# mean-variance trend, helps to see if precision weights are needed
plotSA(fit2, main="Mean variance trend, GSE60740")
